\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in R

### 2014-01-21 papers/jss/article.Rnw
spellcheck.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@806 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

%\paragraph*{Friends don't let friends use CSV!}
Data analysts and researchers often use character seperated text formats such
as \texttt{CSV} \citep{shafranovich2005common} to export and import
data. However, anyone who has ever used \texttt{CSV} files will have noticed
that this method has many limitations: it is restricted to tabular data,
lacks type-safety, and has limited precision for numeric values.  Moreover,
ambiguities in the format itself frequently cause problems.  For example,
conventions on which characters is used as seperator or decimal point vary by
country.  \emph{Extensible Markup Language} (\texttt{XML}) is another
well-established and widely-supported format with the ability to define just
about any arbitrarily complex schema \citep{nolan2013xml}. However, it pays
for this complexity with comparatively large and verbose messages, and added
complexitiy at the parsing side (which are somewhat mitigated by the
availability of mature libraries and parsers). Because \texttt{XML} is text
based and has no native notion of numeric types or arrays, it usually not a
very practical format to store numeric datasets as they appear in statistical
applications.
%
A more modern, widely used format is \emph{JavaScript Object
  Notation} (\texttt{JSON}), which is derived from the object literals of
\proglang{JavaScript}, and used increasingly on the world wide web. \texttt{JSON} natively
supports arrays and distinguishes 4 primitive types: numbers, strings,
booleans and null. However, as it too is a text-based format, numbers are
stored as human-readable decimal notation which is inefficient and
leads to loss of type (double versus integer) and precision. Several R packages
implement functions to parse and generate \texttt{JSON} data from R
objects \citep{rjson,RJSONIO}.

@@

In order to accomplish this, \pkg{RProtoBuf} uses the same catch-all \texttt{proto}
schema used by \pkg{RHIPE} for exchanging R data with Hadoop \citep{rhipe}. This 
schema, which we will refer to as \texttt{rexp.proto}, is printed in appendix 
\ref{rexp.proto}. The Protocol Buffer messages generated by \pkg{RProtoBuf} and
\pkg{RHIPE} are naturally compatible between the two systems because they use the 
same schema. This shows the power of using a schema based cross-platform format such
as Protocol Buffers: interoperability is archieved without effort or close coordination.

>>>>

%\paragraph*{Friends don't let friends use CSV!}
Data analysts and researchers often use character separated text formats such
as \texttt{CSV} \citep{shafranovich2005common} to export and import
data. However, anyone who has ever used \texttt{CSV} files will have noticed
that this method has many limitations: it is restricted to tabular data,
lacks type-safety, and has limited precision for numeric values.  Moreover,
ambiguities in the format itself frequently cause problems.  For example,
conventions on which characters is used as separator or decimal point vary by
country.  \emph{Extensible Markup Language} (\texttt{XML}) is another
well-established and widely-supported format with the ability to define just
about any arbitrarily complex schema \citep{nolan2013xml}. However, it pays
for this complexity with comparatively large and verbose messages, and added
complexity at the parsing side (which are somewhat mitigated by the
availability of mature libraries and parsers). Because \texttt{XML} is text
based and has no native notion of numeric types or arrays, it usually not a
very practical format to store numeric datasets as they appear in statistical
applications.
%
A more modern, widely used format is \emph{JavaScript Object
  Notation} (\texttt{JSON}), which is derived from the object literals of
\proglang{JavaScript}, and used increasingly on the world wide web. \texttt{JSON} natively
supports arrays and distinguishes 4 primitive types: numbers, strings,
booleans and null. However, as it too is a text-based format, numbers are
stored as human-readable decimal notation which is inefficient and
leads to loss of type (double versus integer) and precision. Several R packages
implement functions to parse and generate \texttt{JSON} data from R
objects \citep{rjson,RJSONIO}.

@@

In order to accomplish this, \pkg{RProtoBuf} uses the same catch-all \texttt{proto}
schema used by \pkg{RHIPE} for exchanging R data with Hadoop \citep{rhipe}. This 
schema, which we will refer to as \texttt{rexp.proto}, is printed in appendix 
\ref{rexp.proto}. The Protocol Buffer messages generated by \pkg{RProtoBuf} and
\pkg{RHIPE} are naturally compatible between the two systems because they use the 
same schema. This shows the power of using a schema based cross-platform format such
as Protocol Buffers: interoperability is achieved without effort or close coordination.
